\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath,amssymb}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[hyphens]{url}
\usepackage{microtype}
\usepackage[hidelinks,hypertexnames=false]{hyperref}
\usepackage{caption}
\captionsetup{skip=4pt}
\setlength{\textfloatsep}{10pt plus 2pt minus 2pt}
\setlength{\floatsep}{8pt plus 2pt minus 2pt}
\setlength{\bibsep}{1pt plus 0.2ex}
\emergencystretch=2em
\title{Institutional Robustness Under Capability Escalation in Sequential Commons}
\author{Ameer Alhashemi}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Advanced AI systems are increasingly expected to operate in plural settings rather than as a single monolithic agent. That shift changes the alignment problem from supervising one system to designing institutions that remain robust when many interacting systems face escalating strategic pressure. We study that question in sequential commons. Fishery Commons is retained as a precursor signal-design study showing that central interventions can be ranked under repeated adversarial strategy injection, with adaptive quotas emerging as the strongest overall top-down signal in the medium tier. Harvest Commons becomes the main architecture study. A restored four-condition pilot shows that bottom-up-only governance is distinguishable from both no governance and top-down-only control, but remains ecologically weaker than architectures with central intervention. A high-power confirmatory matrix then shows that hybrid governance ranks first in all eight decision cells under a stronger search-over-mutations attacker. Relative to top-down-only control, hybrid improves patch health in all eight cells, is non-worse on neighborhood overharvest in six of eight, and incurs a moderate mean welfare cost. A capability-ladder analysis over random, mutation, adversarial-heuristic, and search-based attackers shows that the hybrid-versus-top-down comparison is the more sensitive part of the ordering, whereas both governed architectures retain a stable ecological advantage over local-only control. Welfare-incidence logging further shows that the incremental cost of moving from top-down-only to hybrid governance is modest relative to the much larger penalties associated with local-only governance. These results support a benchmark-level claim: in stylized sequential commons, governance architecture can be evaluated as institutional robustness under capability escalation, and multi-scale governance can outperform central-only control without eliminating trade-offs.
\end{abstract}

\input{sections/introduction}
\input{sections/methods}
\input{sections/results}
\input{sections/discussion}

{\small
\bibliographystyle{plainnat}
\bibliography{references}
}

\end{document}
