\section{Discussion}

This follow-up refines the contribution of the existing benchmark rather than broadening it indiscriminately. The earlier two-study paper already showed that governance can be evaluated under repeated adversarial strategy injection in sequential commons, that central signals can be ranked in a minimal setting, and that a narrower hybrid-versus-top-down comparison remains informative under a stronger attacker. The Harvest-led follow-up makes the architecture question explicit. Once the full architecture spectrum is restored, the benchmark no longer asks only whether governance matters. It asks what central-only, local-only, and multi-scale governance each contribute under rising strategic pressure.

The resulting picture is more specific than a generic claim that ``hybrid is best.'' Local-only coordination is not redundant with no governance; it improves some ecological outcomes relative to an unmanaged population. But it remains too weak to preserve the commons when adversarial pressure intensifies. Central control restores ecological stability, and hybrid governance then improves on central-only control enough to win every confirmatory decision cell. The main gain is ecological and behavioral rather than catastrophic. Garden failure is mostly saturated in the confirmatory rows, so the architecture result is carried by patch health, overharvest, and the incidence of restraint rather than by binary collapse alone.

The welfare analysis is equally important to the interpretation. Assessment of governance performance requires explicit attention to the cost of robust control. The incidence summaries show that the large welfare penalty in Harvest Commons is associated with moving away from local-only governance toward architectures that successfully restrain extraction. By contrast, the incremental welfare cost of moving from top-down-only to hybrid governance is moderate. This matters for the institutional reading of the result. It suggests that the benchmark is not simply identifying the most restrictive architecture. It is distinguishing between governance arrangements that differ in how they trade ecological protection against local burden.

The capability ladder further refines that interpretation. It shows that the strongest qualitative split in the benchmark is between architectures with central intervention and local-only governance. That ecological split survives the full non-LLM ladder. The more delicate question is whether hybrid governance preserves an advantage over top-down-only control as attacker strength rises. Here the answer is conditional rather than absolute: the hybrid margin often survives into the stronger rungs, but it is the part of the ordering most sensitive to attacker family, and its welfare threshold can be crossed earlier than its ecological edge disappears. This is still substantively informative. It identifies not only which architecture ranks first in the strongest confirmatory matrix, but also which comparison is robust and which comparison remains contestable as strategic pressure escalates.

At the same time, the scope of the claim remains deliberately narrow. The paper does not study all possible architectures, all possible attacker models, or all relevant notions of alignment. Harvest LLM remains outside the main evidence chain because it has not cleared the reliability gate. The environments remain stylized, and the benchmark should be read as a tool for controlled comparison rather than as a direct model of deployment. Within those bounds, however, the follow-up supports a stronger empirical claim than the earlier draft: governance architecture in sequential commons can be evaluated as institutional robustness under capability escalation, and multi-scale governance can outperform central-only control without eliminating trade-offs.
