\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath,amssymb}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[hyphens]{url}
\usepackage{microtype}
\usepackage[hidelinks,hypertexnames=false]{hyperref}
\usepackage{float}
\usepackage{caption}
\captionsetup{skip=4pt}
\setlength{\textfloatsep}{10pt plus 2pt minus 2pt}
\setlength{\floatsep}{8pt plus 2pt minus 2pt}
\setlength{\intextsep}{8pt plus 2pt minus 2pt}
\setlength{\bibsep}{1pt plus 0.2ex}
\emergencystretch=2em
\title{Governance Under Adversarial Strategy Injection in Sequential Commons}
\author{Ameer Alhashemi}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We study governance under adversarial strategy injection in sequential commons through two linked studies. The first study uses a minimal Fishery Commons setting to compare central governance signals under repeated strategy turnover and held-out regime shifts. A matched mutation-versus-live-LLM baseline shows that monitoring with sanctions reduces held-out collapse relative to no governance for both injector types, a broader top-down sweep over partner mixes and pressure levels identifies adaptive quotas as the strongest overall signal, and a learned-policy self-play validation in the medium tier recovers the same ordering: adaptive quotas beat no governance on held-out collapse, stock, and welfare while also outperforming the other governed conditions. The second study upgrades Harvest Commons to the same invasion logic and compares governance architectures rather than only signal choices. After a fixed-composition pilot and a lighter non-LLM invasion pass established that hybrid governance was promising, a stronger search-over-mutations stress test became the main Harvest result. In that high-power matrix, the hybrid condition wins all eight decision cells against top-down-only control, improving patch health and usually reducing neighborhood overharvest, though at a moderate average welfare cost. A narrower medium-tier PPO validation points in the same direction: hybrid self-play improves patch health and lowers local aggression and neighborhood overharvest relative to top-down-only control, with only a small welfare cost. The combined result is a two-level evaluation protocol: first identify effective central governance signals in a controlled commons, then test how broader governance architectures behave in a second substrate under the same turnover pressure.
\end{abstract}

\input{sections/introduction}
\input{sections/methods}
\input{sections/theory}
\input{sections/results}
\input{sections/discussion}
\input{sections/limitations}
\input{sections/appendix}

{\small
\bibliographystyle{plainnat}
\bibliography{references}
}

\end{document}
