\section{Limitations}

The two studies answer adjacent questions, not identical ones. The first study is a controlled signal-ranking study in Fishery Commons, while the second study is a governance-architecture extension in Harvest Commons. The second study now shares the same invasion loop and replacement logic as the first, which fixes the earlier asymmetry, but it still should not be read as a strict replication in another substrate because the governance question itself is broader.

The LLM comparison currently appears only in the Fishery Commons study as main paper evidence. This is enough to show that minimal central governance survives both mutation and live JSON-based strategy generation, but it does not yet establish the same result for Harvest Commons. A live Harvest LLM path has now been implemented and can pass a narrow reliability screen, but the matched narrow Harvest pilot still failed the manuscript's evidence gate because schema completeness remained too uneven: the mean effective-LLM fraction stayed at \(0.7311\) and unrepaired fallback averaged \(0.2689\), with missing keys as the dominant failure mode. For that reason, Harvest LLM remains capability-only evidence and the main Study 2 claim stays anchored in the stronger non-LLM search-over-mutations result.

The learned-policy extensions are still narrow. In Fishery Commons, the PPO validation is limited to the medium tier and fixed governance conditions. In Harvest Commons, a medium-tier PPO self-play validation now supports the `hybrid > top\_down\_only` direction, but it still uses fixed governance conditions, shaped rewards, and held-out evaluation in only one tier. These checks are enough to show that the leading Fishery and Harvest rankings are not artifacts of one hand-specified policy family, but they are not yet learned-policy invasion benchmarks and they do not yet replace the main mutation or search-over-mutations evidence.

Hard settings remain useful mainly as stress evidence. In fishery, once collapse saturates at 1.0 and survival at 0.0 across conditions, governance separation becomes non-identifiable under floor effects. In Harvest Commons, the harder tiers remain informative, but garden failure itself does not separate conditions in the high-power rows; the real signal comes from patch health, welfare, and local behavioral metrics.

Finally, both substrates remain intentionally stylized. That is a strength for identification, but it limits external realism. Market feedbacks, richer heterogeneity, richer language interaction, and explicit institutional bargaining are natural next additions once the current governance comparisons are frozen.
