\section{Methods}

Both studies use the same core invasion protocol. A population of strategies is evaluated over generations, weaker strategies are removed, and new strategies are injected through a specified generator. Selection is performed in train conditions and reported robustness is measured in held-out test conditions. This separates adaptation pressure from generalization assessment and makes governance comparisons about out-of-distribution stability rather than only in-sample performance.

\subsection{First study: Fishery Commons and central governance signals}

The first study uses a minimal renewable fishery commons with a shared stock, per-agent harvest actions, regeneration, and collapse dynamics. The main invasion runs use interpretable threshold strategies, and governance changes extraction pressure directly through monitoring, quota clipping, sanctions, adaptive quotas, or temporary closures. The environment logs requested harvest, realized harvest, quota-clipped harvest, repeat-offender rate, closure activity, and stock-recovery lag so that governance can be analyzed mechanistically rather than only through final collapse outcomes.

The Fishery Commons study has two invasion layers plus one learned-policy validation. The first invasion layer is a matched baseline that compares no governance, monitoring alone, and monitoring with sanctions across two injector types: mutation and live JSON-based LLM injection through Ollama with the `qwen2.5:3b-instruct` model. This baseline is run on the easy and medium fishery tiers. The second invasion layer is a broader top-down sweep, restricted to the medium tier, that compares no governance, monitoring with sanctions, adaptive quotas, and temporary closures under three non-LLM injector families: mutation, an adversarial heuristic injector, and search over mutations. The sweep also varies partner composition between balanced and adversarial-heavy populations, and it varies adversarial pressure between 0.3 and 0.5. This layer identifies which central governance signals remain effective once the setting becomes more hostile.

The learned-policy validation asks a narrower question: does the resulting governance ranking survive outside the hand-specified threshold family? For this check, agents share a parameter-shared PPO policy with a small multilayer perceptron, observe normalized stock and normalized timestep, and choose among 11 harvest bins. Training uses the same fishery train regime as the matched baseline (`regen\_rate = 2.0`, `obs\_noise\_std = 6.0`) for 250{,}000 environment steps per seed. Final checkpoint selection uses train-regime evaluation only, and held-out reporting uses the fixed `medium\_v1` benchmark pack. A direct five-seed confirmatory comparison evaluates no governance against adaptive quotas, and a matched five-seed governance follow-up compares monitoring with sanctions, adaptive quotas, and temporary closures under the same PPO protocol. This evaluation serves as a policy-family robustness check on the top-down ranking rather than as an additional invasion matrix. The shared Fishery invasion defaults for the matched baseline are 15 generations, 32 train and 32 test seeds per generation, replacement fraction 0.2, and adversarial pressure 0.3.

\subsection{Second study: Harvest Commons and governance architecture}

The second study introduces Harvest Commons, a second renewable commons substrate built around local patches rather than a single global stock. It uses the same invasion logic as the first study rather than only a fixed-composition matrix. A population of six parameterized Harvest strategies is evaluated, ranked, and partially replaced over generations. The evolving strategies control local harvest fractions, restraint signals, credit requests, credit offers, reciprocity weights, and cap-compliance margins. Communication and side-payments remain condition-controlled by the environment, so the same evolving strategy family can be evaluated under different governance architectures.

Across the broader Harvest study, four governance conditions are defined: none, top-down only, bottom-up only, and hybrid. Top-down only uses a fixed government agent that observes patch stress and local aggressive neighborhoods, then applies targeted caps. Bottom-up only enables local communication and side-payments without central intervention. Hybrid combines the fixed government agent with local communication and credit transfers. The earliest Harvest evidence is a fixed-composition pilot over that broader architecture set, retained only as precursor evidence. A subsequent invasion pass with lighter non-LLM injectors suggested that hybrid governance warranted closer analysis. The principal Harvest comparison in the manuscript narrows to a stronger search-over-mutations comparison between top-down-only and hybrid, where the invader repeatedly proposes the most damaging child among multiple mutated candidates.

The strengthened Harvest cycle begins with a pilot matrix over the medium and hard tiers, the balanced and adversarial-heavy partner mixes, top-down only versus hybrid governance, mutation versus search over mutations, and adversarial pressures 0.3 and 0.5. The decision-critical Stage C matrix then fixes the injector to search over mutations and reruns those same medium and hard, balanced and adversarial-heavy, top-down and hybrid cells at higher power. Each decision cell is one `(tier, partner mix, pressure)` row in that Stage C matrix, and each cell is run for 15 generations with 32 train and 32 test seeds per generation. In addition to patch health and welfare, the Harvest Commons study reports local mechanism metrics including local aggression, neighborhood overharvest, capped-action rate, targeted-agent fraction, and prevented harvest. These metrics are necessary because a governance architecture can change \emph{where} exploitation occurs, not only the global average outcome.

The second study also includes a narrow learned-policy validation. For this check, agents share a parameter-shared PPO policy with separate communication and action heads. Communication is chosen from local observations, while harvest and credit-offer decisions condition on local observations plus inbox summaries from neighboring agents. The validation is restricted to the `medium\_h1` tier and compares top-down-only against hybrid governance over five training seeds, each run for 100{,}000 environment steps. Training reward combines welfare with a patch-health bonus and mild penalties on local aggression and neighborhood overharvest. Held-out reporting uses the fixed `medium\_h1` benchmark pack. This validation is reported as supporting evidence for the architecture comparison rather than as a replacement for the main Stage C invasion matrix.

\subsection{Statistical reporting}

The primary matched-baseline outcome in the Fishery Commons study is held-out collapse reduction, reported as the difference between no governance and monitoring with sanctions. Supporting metrics are held-out stock, time-to-collapse, and per-regime survival. Confidence intervals for the matched baseline are run-level bootstrap intervals with 5{,}000 resamples. The broader Fishery Commons sweep uses the same run summaries but is interpreted as a ranking study: adaptive quota, temporary closure, and monitoring with sanctions are compared on collapse first and welfare second. The learned-policy validation reports the same held-out collapse, stock, welfare, and per-regime survival summaries over five independent PPO training seeds.

The primary comparison in the Harvest Commons study is the paired difference between hybrid and top-down-only governance at matched run identifiers within each tier, partner mix, and adversarial-pressure cell. This comparison isolates whether local communication and credit sharing add anything once central intervention is already present. Confidence intervals are again estimated by bootstrap over paired deltas. The manuscript also reports best-condition counts over the Stage C decision cells. Those counts use the same ranking rule as the archived summary artifacts: lower garden failure first, higher patch health second, and higher welfare third. Accordingly, references to a condition ``winning'' a decision cell denote first rank under that lexicographic rule rather than superiority on every reported metric. In the high-power Stage C matrix, garden failure is never the discriminating outcome, so the main Harvest Commons evidence is carried by patch health, welfare, local aggression, and neighborhood overharvest. The supporting Harvest PPO validation reports held-out means over five independent training seeds in `medium\_h1`; because garden failure is saturated there as well, interpretation again focuses on patch health, welfare, local aggression, and neighborhood overharvest.

Formal invasion dynamics and the saturation identifiability condition are provided in supplementary materials and source files. In the empirical pipeline, the reported quantities are estimated from run-level summaries and archived result artifacts to avoid post-hoc threshold drift.
